One goal, two paths: How dynamic capabilities enable competitive advantage through behavioral and non-behavioral objects of change

### 2020-01-22 article4_jbr.rnw
Finalized for inclusion in dissertation

\section{Introduction}

<<loading, results="asis", echo=FALSE >>=
setwd("/Users/larshovdanmolden/Documents/git/article4")
library(xtable)
library(semPLS)
#load("./data/resart4.RData")
r <- main$resmat
rm  <- main$resmatmed
l  <- main$loadings
ic <- data.frame(i[,c(1,4)]);colnames(ic) <- c("b","t")
ic$rn  <- 1:nrow(ic)
bb <- main$models$both
bm <- main$models$mmed
br <- main$models$rmed
#foolars=0.99
@

@@

\section{Theoretical background and hypotheses}

@@

\subsection{Conducting and coaching at Microsoft}\label{sec:case}

@@

\subsection{Capabilities conductor}

@@

\subsection{Cognitive coach}

@@

\subsection{Coaching conductor}

@@

\begin{figure}
  \centering
  \captionsetup{width=0.5\linewidth}
  \input{./figure/fullmod.tex}
  \captionof{figure}{The resource and cognitive mechanisms of dynamic capability}
  \label{fig:fig1}
\end{figure}

\section{Data and methods}

@@

\subsection{Variable construction}

@@

\subsection{Control variables}

@@

\begin{table}[t]
  \caption{Descriptive statistics and correlation between key constructs and control
  variables.}
\label{tab:cor}
  <<corchild, child='./figure/cortab.rnw',results="asis">>=
@

@@

\subsection{Estimation and empirical model}

@@

\section{Results}
The main results of our analysis are depicted both in the graphical path diagrams (figure
\ref{fig:res}) and the regression table in table \ref{tab:res}. The use of PLS-SEM has
increased substantially in the realm of marketing \citep{hult2018} and strategic management
\citep{Hair2012,Valaei2017} and is the topic of special issues in leading journals \citep{Hair2020} (e.g. Journal of
Business Research call for papers in 2020). In this tradition we aim to present thorough
analysis of our models and largely follows the recommendations made in recent
contributions \citep{Hair2019}.

@@

\begin{table}[t]
  \caption{PLS-SEM results for key constructs}
\label{tab:res}
  \scriptsize
  <<child-demo, child='./figure/restab.rnw'>>=
@
   \begin{tablenotes}
      \tiny
      \item Standard-errors estimated using bootstrapping with 10 000 simulations.
      \item Firm level control variables include firm size, firm age,
        profit margin, firm assets.
    \end{tablenotes}
  \end{table}

\subsection{Robustness and assumption validation}
The results are subject to some robustness checks not directly reported in the paper,
albeit a summary is available in the appendix. First, controlling for functional form
using a Ramsay RESET test we find support for a linear relationship between our
constructs. Second, we ran the a Variance Inflation Test (VIF) to control for
multicolinearity. None of the VIF values exceeded the suggested level indicating
multicolinearity problems. Third, we ran two separate models where we (1) included other
measures of $DC$ (using the original measure of $DC_0$), and (2) added more control
variables to which decreased the sample size (from $N=238$ to $N=205$).

@@

 <<pathchild, child='./figure/medanalysisall.rnw',results="asis">>=
@

\section{Discussion}
The aim of this paper is to theorize about how non-behavioral objects of change work as a
mechanism through which dynamic capabilities influence competitive advantage. Our results
indicate that both behavioral (\emph{operating capabilities}) and non-behavioral
(\emph{strategic cognition}) mediates the effect from dynamic capabilities on competitive
advantage. The first mechanism, dubbed 'conducting' in reference to the well established
terminology of 'resource orchestration' \citep{Sirmon2011,Helfat2007}, shows a robust and
strong function as a mediator in line with our first hypothesis. When firms exhibit strong
dynamic capabilities they are able to adapt their operating capabilites to stay
competitive. We hypothesized that this relationship would be positive and significant. Our
results indicated exactly this relationship and thus lending support to hour hypothesis.
In our example context of Microsoft Nadella did several investments in capabilities and
resources as a response to Microsofts dynamic capabilities execution. For example, he
acquired important capabilities from M\&A efforts, and upgraded the competencies regarding
cloud business and big data \citep{Nadella2017}. This should not be surprising given the
vast literature on indirect effects of dynamic capabilities
\citep{Schilke2018,Eriksson2014} supported by empirical and theoretical research
\citep{Pezeshkan2016b,Fainshmidt2016a,Protogerou2012,Karimi2015}. However, it adds to this
ongoing effort of investigating mechanisms through which dynamic capabilities work which
is called for in the literature \citep{Schilke2018}.

@@

\section{Concluding remarks}

@@

\section{Bibliography}

@@

\section{Appendix}
\newpage

  \begin{table}\label{tab:items}
    \caption{Core constructs and their measurement items. Factor
      loadings in parantheses}
    <<itemchild, child='./figure/itemtab.rnw',results="asis">>=
    @
    \begin{tablenotes}
      \tiny
      \item Dillon-Goldstein Rho is recommended as reliability measure
        for SEM-PLS \citep{Hair2019}.
    \end{tablenotes}
  \end{table}

>>>>

\section*{Introduction}

<<loading, results="asis", echo=FALSE >>=
setwd("/Users/larshovdanmolden/Documents/git/article4")
library(xtable)
library(semPLS)
load("./data/resart4.RData")
r <- main$resmat
rm  <- main$resmatmed
l  <- main$loadings
ic <- data.frame(i[,c(1,4)]);colnames(ic) <- c("b","t")
ic$rn  <- 1:nrow(ic)
bb <- main$models$both
bm <- main$models$mmed
br <- main$models$rmed
#foolars=0.99
@

@@

\section*{Theoretical background and hypotheses}

@@

\subsection*{Conducting and coaching at Microsoft}\label{sec:case}

@@

\subsection*{Capabilities conductor}

@@

\subsection*{Cognitive coach}

@@

\subsection*{Coaching conductor}

@@

\begin{figure}
  \centering
  \captionsetup{width=0.5\linewidth}
  \input{/Users/larshovdanmolden/Documents/git/article4/figure/fullmod.tex}
  \captionof{figure}{The resource and cognitive mechanisms of dynamic capability}
  \label{fig:fig1}
\end{figure}

\section*{Data and methods}

@@

\subsection*{Variable construction}

@@

\subsection*{Control variables}

@@

\begin{table}[t]
  \caption{Descriptive statistics and correlation between key constructs and control
  variables.}
\label{tab:cor}
  <<corchild, child='/Users/larshovdanmolden/Documents/git/article4/figure/cortab.rnw',results="asis">>=
@

@@

\subsection*{Estimation and empirical model}

@@

\section*{Results}
The main results of our analysis are depicted both in the graphical path diagrams (figure
\ref{fig:res}) and the regression table in table \ref{tab:res}. The use of PLS-SEM has
increased substantially in the realm of marketing \citep{hult2018} and strategic management
\citep{Hair2012,Valaei2017} and is the topic of special issues in leading journals \citep{Hair2020} (e.g. Journal of
Business Research call for papers in 2020). In this tradition we aim to present thorough
analysis of our models and largely follows the recommendations made in recent
contributions \citep{Hair2019}.

@@

\begin{table}[t]
  \caption{PLS-SEM results for key constructs}
\label{tab:res}
  \scriptsize
  <<child-demo, child='/Users/larshovdanmolden/Documents/git/article4/figure/restab.rnw'>>=
@
   \begin{tablenotes}
      \tiny
      \item Standard-errors estimated using bootstrapping with 10 000 simulations.
      \item Firm level control variables include firm size, firm age,
        profit margin, firm assets.
    \end{tablenotes}
  \end{table}

\subsection*{Robustness and assumption validation}
The results are subject to some robustness checks not directly reported in the paper,
albeit a summary is available in the appendix. First, controlling for functional form
using a Ramsay RESET test we find support for a linear relationship between our
constructs. Second, we ran the a Variance Inflation Test (VIF) to control for
multicolinearity. None of the VIF values exceeded the suggested level indicating
multicolinearity problems. Third, we ran two separate models where we (1) included other
measures of $DC$ (using the original measure of $DC_0$), and (2) added more control
variables to which decreased the sample size (from $N=238$ to $N=205$).

@@

 <<pathchild, child='/Users/larshovdanmolden/Documents/git/article4/figure/medanalysisall.rnw',results="asis">>=
@

\section*{Discussion}
The aim of this paper is to theorize about how non-behavioral objects of change work as a
mechanism through which dynamic capabilities influence competitive advantage. Our results
indicate that both behavioral (\emph{operating capabilities}) and non-behavioral
(\emph{strategic cognition}) mediates the effect from dynamic capabilities on competitive
advantage. The first mechanism, dubbed 'conducting' in reference to the well established
terminology of 'resource orchestration' \citep{Sirmon2011,Helfat2007}, shows a robust and
strong function as a mediator in line with our first hypothesis. When firms exhibit strong
dynamic capabilities they are able to adapt their operating capabilites to stay
competitive. We hypothesized that this relationship would be positive and significant. Our
results indicated exactly this relationship and thus lending support to hour hypothesis.
In our example context of Microsoft Nadella did several investments in capabilities and
resources as a response to Microsofts dynamic capabilities execution. For example, he
acquired important capabilities from M\&A efforts, and upgraded the competencies regarding
cloud business and big data \citep{Nadella2017}. This should not be surprising given the
vast literature on indirect effects of dynamic capabilities
\citep{Schilke2018,Eriksson2014} supported by empirical and theoretical research
\citep{Pezeshkan2016b,Fainshmidt2016a,Protogerou2012,Karimi2015}. However, it adds to this
ongoing effort of investigating mechanisms through which dynamic capabilities work which
is called for in the literature \citep{Schilke2018}.

@@

\section*{Concluding remarks}

@@

\section*{Bibliography}

@@

\section*{Appendix}
\newpage

  \begin{table}\label{tab:items}
    \caption{Core constructs and their measurement items. Factor
      loadings in parantheses}
    <<itemchild, child='/Users/larshovdanmolden/Documents/git/article4/figure/itemtab.rnw',results="asis">>=
    @
    \begin{tablenotes}
      \tiny
      \item Dillon-Goldstein Rho is recommended as reliability measure
        for SEM-PLS \citep{Hair2019}.
    \end{tablenotes}
  \end{table}
